Diagnostics Week of June 28th

### 2020-07-02 output_scripts/Diagnostics.Rmd
add the c-hp,dc-hp and c-mmhp plots for a given pair and window. need to check c-mmhp

theme_set(theme_minimal())
```

@@

@@

```{r}
no_segments <- 500

current_cohort <- 5
current_i <- 8
current_j <- 2
current_win <- 13
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

@@

```{r plot_functions, include=FALSE}
drawRealIntensityPaper<-function(t,object=NULL,simulation=NULL,yupper=10,add=FALSE,color=1,line.width=1,
                                 title_name="",title.cex=9, y.ratio=0, min.y=0, min.x=0, max.x=NULL, 
                                 annotate.time=FALSE, event.cex=1, box.type="l", box.col="gray40", title.line=1){
  # input object: the parameter list used for generating mmhp
  #       simulation: simulation result from simulate.mmhp 
  if(t[1]>0){
    t <- c(0,t)
  }
  if(is.null(max.x)){
    max.x <- tail(t,1)
  }
  if(is.null(object)){
    termination <- tail(t,1)
    n <- length(t)
    plot(0,0,xlim=c(min.x,max.x),ylim=c(min.y,yupper*(1+y.ratio)),
         type="n",xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
    box(lty ='solid', col=box.col, lwd=0.7, bty=box.type)
    title(title_name,line=title.line,cex.main=title.cex)
    #left axes
    #axis(2,at=seq(0,yupper,5),labels=FALSE,cex.lab=2,lwd=0,lwd.ticks=1) 
    #text(-termination/50, y=seq(0,yupper,5), labels=seq(0,yupper,5), cex=1.6, srt=0, xpd=TRUE)
    #bottom axes
    # events
    points(t[-1],rep(min.y,n-1),cex=event.cex,pch=16,col="black")
  }else{
    state <- simulation$z
    state_time <- simulation$x
    if(tail(state_time,1)<tail(t,1)){
      state_time <- c(state_time,tail(t,1))
      state <- c(state,3-tail(state,1))
    }
    lambda0 <- object$lambda0
    lambda1 <- object$lambda1
    alpha <- object$alpha
    beta <- object$beta
    termination <- tail(t,1)
    n <- length(t)
    m <- length(state)
    if(is.null(min.y)){
      min.y <- -lambda0
    }
    if(is.null(max.x)){
      max.x <- termination
    }
    #plot(1,4,xlim=c(0.27,23),ylim=c(-0.5,9), type="n",xlab="Time",ylab="",xaxt="n",yaxt="n",cex.lab=1.8,bty="n")

@@

drawHawkesIntensityPaper<-function(lambda0,alpha,beta,events,color=1,line.width=2, termination=NULL){
  # input object: parameters for Hawkes process, include lambda0, alpha, beta 
  #       events: vector of event happening time
  horizon <- tail(events,1)
  events <- events[events>0]
  N <- length(events)

  if(N==1){ # only one event condition
    segments(x0=0, x1=events[1], y0=lambda0, col=color,lwd=line.width)
    segments(x0=events[1], y0=lambda0, y1=lambda0+alpha, col=color,lwd=line.width)
    if(horizon > tail(events,1)){
      lambda.n <- function(s) lambda0+alpha*sum(exp(-beta*(s-events[1])))
      new.lambda.n<-Vectorize(lambda.n)
      curve(new.lambda.n, from=tail(events,1), to=horizon, add=TRUE, col=color,lwd=line.width)
    }
    if( !is.null(title) ) title(main=title)
  }else{
    segments(x0=0, x1=events[1], y0=lambda0, col=color,lwd=line.width)
    segments(x0=events[1], y0=lambda0, y1=lambda0+alpha, col=color,lwd=line.width)

@@

```{r, fig.width=10}
#par(mfrow=c(2,5), oma=c(1,1,1,1), mar=c(2,3,2,1), mgp=c(0.8,0,0))
#layout(matrix(1:10,nrow=2,byrow=TRUE),widths=rep(1,5),height=c(3,3))
anno.cex <- 1.3
title.cex <- 2.2
title.line <- 0.5
y.upper <- 20
## pp
lambda.pp <- 1/mean(diff(c(0,current_event_time,observe_period)))
drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="Poisson", title.cex=title.cex, max.x=observe_period, box.col="black")
abline(h=lambda.pp,lwd=2)
mtext(side=1, text="Time", line=1.6, cex=anno.cex)
mtext(side=2, text="Intensity", line=1.2, cex=anno.cex)

## Hawkes
hawkes.par<-optim(par=rep(1,3), fn=uniHawkesNegLogLik, method="L-BFGS-B", 
                  t=current_event_time, termination = observe_period, lower=c(0,1,0))

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="Hawkes", title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par$par[1],
                         alpha=hawkes.par$par[2],
                         beta=hawkes.par$par[3],
                         events=current_event_time, termination=observe_period)
mtext(side=1, text="Time", line=1.6, cex=anno.cex)
mtext(side=2, text="Intensity", line=1.2, cex=anno.cex)
mtext(side=1, text="Time", line=1.6, cex=anno.cex)
mtext(side=2, text="Intensity", line=1.2, cex=anno.cex)

@@

```{r load_mmhp_data, cache=TRUE}
data_path <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_june_26/output//"

## MMHP
load(paste(data_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(data_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1),
                 lambda0= (mean(sim_cohort_mmhp$gamma[,current_i])+
                             mean(sim_cohort_mmhp$zeta[,current_j])),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

@@

@@

```{r}
#y.upper <- 10
#title.line <- 1
#title.cex <- 0.1
drawRealIntensityPaper(current_event_time, object=mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,mmhp.est.latent$x.hat),
                                       z=mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")
mtext(side=1, text="Time", line=1.6, cex=anno.cex)
mtext(side=2, text="Intensity", line=1.2, cex=anno.cex)

@@

```{r}
anno.cex <- 1
#### Hawkes
Lambda.hawkes.test <- Lambda_hawkes_matrix[current_i,current_j][[1]]

>>>>

theme_set(theme_minimal())

other_path  <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_null_windows/"
old_data_path  <- "C:/Users/owenw/Google Drive/Tian/Current_Projects/MMHP_Latent/output_null_windows/"

@@

for(i in 1:mice_number){
  print(i)
  for(j in 1:mice_number){
    pair <- which(unique_pairs_df$initiator==i&unique_pairs_df$recipient==j)
    if(length(pair)>0&(i!=j)){
      current_window_vec <- unique_pairs_df$observe[[pair]]
      all_rescaled_interevent <- numeric(0)
      for(cur in c(1:length(current_window_vec))){ 
        cur_win <- current_window_vec[cur]
        current_event_time <- return_df[return_df$initiator==i&
                                        return_df$recipient==j&
                                        return_df$observe.id==cur_win,"event.times"][[1]]
        observe_period <- return_df[return_df$initiator==i&
                                        return_df$recipient==j&
                                        return_df$observe.id==cur_win,"observe.time"]
        #### update this from the stan fit
        hawkes.par <- c(model2_par_matrix$lambda0_matrix[i,j],model2_par_matrix$alpha[i,j],
                        model2_par_matrix$beta[i,j])
        # hawkes.par <- optim(par=c(1,0,100), fn=uniHawkesNegLogLik, method="CG", 
        #                     t=current_event_time, termination = observe_period)
        all_Lambda <- uniHawkesIntegralIntensity(object=list(lambda0=hawkes.par[1],
                                                              alpha=hawkes.par[2],
                                                              beta=hawkes.par[3]),
                                                  events=current_event_time,
                                                  termination = observe_period)
        all_prresidual <- uniHawkesPearsonResidual(object=list(lambda0=hawkes.par[1],
                                                              alpha=abs(hawkes.par[2]),
                                                              beta=hawkes.par[3]),
                                                  events=current_event_time,
                                                  termination = observe_period)

@@

```{r select cohort pair window}
no_segments <- 500

current_cohort <- 4#5
current_i <- 2 #8
current_j <- 8 #2
current_win <- 42
clean_data <- cleanData(full_data[[cohort_names[current_cohort]]])
return_df <- cleanObservationPeriod(current_cohort, clean_data)
unique_pairs_df <- return_df %>% group_by(initiator, recipient) %>%
  dplyr::summarize(count=n(),
            observe=list(observe.id),
            observe.length=list(observe.time),
            no.events=list(no.events))
pair <- which(unique_pairs_df$initiator==current_i&unique_pairs_df$recipient==current_j)
unique_observe_win <- unique(return_df[,c("observe.id","observe.time")])

@@

```{r plot_functions, include=FALSE}
drawRealIntensityPaper <- function(t,
                                   object=NULL,simulation=NULL,
                                   yupper=10,add=FALSE,color=1,line.width=1,
                                 title_name="",title.cex=9, y.ratio=0,
                                 min.y=0, min.x=0, max.x=NULL, 
                                 annotate.time=FALSE,
                                 event.cex=1, box.type="l",
                                 box.col="gray40", title.line=1){
  # input object: the parameter list used for generating mmhp
  #       simulation: simulation result from simulate.mmhp 
  if(t[1]>0){
    t <- c(0,t)
  }
  if(is.null(max.x)){
    max.x <- tail(t,1)
  }
  if(is.null(object)){
    termination <- tail(t,1)
    n <- length(t)
    plot(0,0,xlim=c(min.x,max.x),ylim=c(min.y,yupper*(1+y.ratio)),
         type="n",xlab="",ylab="",xaxt="n",yaxt="n",axes=FALSE)
    box(lty ='solid', col=box.col, lwd=0.7, bty=box.type)
    title(title_name,line=title.line,cex.main=title.cex)
    #left axes
    #axis(2,at=seq(0,yupper,5),labels=FALSE,cex.lab=2,lwd=0,lwd.ticks=1) 
    #text(-termination/50, y=seq(0,yupper,5), labels=seq(0,yupper,5), cex=1.6, srt=0, xpd=TRUE)
    #bottom axes
    # events
    points(t[-1],rep(min.y,n-1),cex=event.cex,pch=16,col="black")
  }else{
    state <- simulation$z
    state_time <- simulation$x
    if(tail(state_time,1)<tail(t,1)){
      state_time <- c(state_time,tail(t,1))
      state <- c(state,3-tail(state,1))
    }
    lambda0 <- object$lambda0
    lambda1 <- object$lambda1
    alpha <- object$alpha
    beta <- object$beta
    termination <- tail(t,1)
    n <- length(t)
    m <- length(state)
    if(is.null(min.y)){
      min.y <- -lambda0
    }
    if(is.null(max.x)){
      max.x <- termination
    }
    #plot(1,4,xlim=c(0.27,23),ylim=c(-0.5,9), type="n",xlab="Time",ylab="",xaxt="n",yaxt="n",cex.lab=1.8,bty="n")

@@

drawHawkesIntensityPaper <- function(lambda0,alpha,beta,
                                     events,color=1,line.width=2, termination=NULL){
  # input object: parameters for Hawkes process, include lambda0, alpha, beta 
  #       events: vector of event happening time
  horizon <- tail(events,1)
  events <- events[events>0]
  N <- length(events)

  if(N==1){ # only one event condition
    segments(x0=0, x1=events[1], y0=lambda0, col=color,lwd=line.width)
    segments(x0=events[1], y0=lambda0, y1=lambda0+alpha, col=color,lwd=line.width)
    if(horizon > tail(events,1)){
      lambda.n <- function(s) lambda0+alpha*sum(exp(-beta*(s-events[1])))
      new.lambda.n<-Vectorize(lambda.n)
      curve(new.lambda.n, from=tail(events,1), to=horizon, add=TRUE, col=color,lwd=line.width)
    }
    #if( !is.null(title) ) title(main=title)
  }else{
    segments(x0=0, x1=events[1], y0=lambda0, col=color,lwd=line.width)
    segments(x0=events[1], y0=lambda0, y1=lambda0+alpha, col=color,lwd=line.width)

@@

```{r plot intensity in a window, fig.width=10}
y.upper <- 20
title.line <- 1
title.cex <- 1
anno.cex <- 1

## Cohort Hawkes ####
# load the parameters for this pair
load(paste(old_data_path, cohort_names[current_cohort],
           "/cohort_hp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

hawkes.par <- c(model1_par_matrix$lambda0_matrix[current_i,current_j],
                model1_par_matrix$alpha[current_i,current_j],
                        model1_par_matrix$beta[current_i,current_j])

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="Cohort Hawkes",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)
mtext(side=1, text="Time", line=1.6, cex=anno.cex)
mtext(side=2, text="Intensity", line=1.2, cex=anno.cex)

#### Degree Corrected Hawkes ####
load(paste(old_data_path, cohort_names[current_cohort],
           "/cohort_dchp_stan_result_", cohort_names[current_cohort],".RData",sep=''))

hawkes.par <- c(model2_par_matrix$lambda0_matrix[current_i,current_j],
                model2_par_matrix$alpha[current_i,current_j],
                model2_par_matrix$beta[current_i,current_j])

drawRealIntensityPaper(current_event_time,
                       yupper=y.upper, color="black",line.width=2,title.line=title.line,
                       title_name="Cohort DC-Hawkes",
                       title.cex=title.cex, max.x=observe_period, box.col="black")
drawHawkesIntensityPaper(lambda0=hawkes.par[1],
                         alpha=hawkes.par[2],
                         beta=hawkes.par[3],
                         events=current_event_time, termination=observe_period)
mtext(side=1, text="Time", line=1.6, cex=anno.cex)
mtext(side=2, text="Intensity", line=1.2, cex=anno.cex)

rm(fit_cohort_hp)
rm(fit_cohort_dchp)
rm(sim_cohort_hp)
rm(sim_cohort_dchp)

@@

```{r draw_mmhp_fit}
#current_cohort <- 2
## Cohort MMHP
load(paste(c_mmhp_path,cohort_names[current_cohort],
             "/cohort_mmhp_stan_result_",cohort_names[current_cohort],
             ".RData",sep=''))
load(paste(c_mmhp_path,cohort_names[current_cohort],
             "/cmmhp_est_zt_",cohort_names[current_cohort],".RData",sep=''))

# then adjust the parameters for c-mmhp below here also

mmhp_par <- list(lambda1=mean(sim_cohort_mmhp$lambda1[,pair]),
                 lambda0=mean(sim_cohort_mmhp$lambda0[,pair]),
                 alpha=mean(sim_cohort_mmhp$alpha[,pair]),
                 beta=mean(sim_cohort_mmhp$beta),
                 q1=mean(sim_cohort_mmhp$q1[,pair]),
                 q2=mean(sim_cohort_mmhp$q2[,pair]))

@@

drawRealIntensityPaper(current_event_time, object=mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,mmhp.est.latent$x.hat),
                                       z=mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="Cohort MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")
mtext(side=1, text="Time", line=1.6, cex=anno.cex)
mtext(side=2, text="Intensity", line=1.2, cex=anno.cex)

rm(interpolation_array_list)
rm(state_array_list)
rm(initial_state_list)
rm(termination_state_list)

@@

```{r, eval=FALSE}
#y.upper <- 10
#title.line <- 1
#title.cex <- 0.1
drawRealIntensityPaper(current_event_time, object=mmhp_par,title.line=title.line,
                       simulation=list(x=c(0,mmhp.est.latent$x.hat),
                                       z=mmhp.est.latent$z.hat),
                       yupper=y.upper, color="black", line.width=2,
                       title_name="MMHP", title.cex=title.cex, max.x=observe_period, box.col="black")
mtext(side=1, text="Time", line=1.6, cex=anno.cex)
mtext(side=2, text="Intensity", line=1.2, cex=anno.cex)

@@

```{r, eval=FALSE}
anno.cex <- 1
#### Hawkes
Lambda.hawkes.test <- Lambda_hawkes_matrix[current_i,current_j][[1]]
